%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%  esrel2020-paper.tex   :   08/11/2019 			                     %%
%%  Text file to use with rps-esrel2020. written in Latex2e.             %%
%%  The content, structure, format and layout of this style file is the  %%
%%  property of Research Publishing Services                             %%
%%  Copyright (c) 2011-2018 Research Publishing Services,                %%
%%  All rights are reserved.                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\documentclass[twocolumn]{rps-esrl2020}
\usepackage{lipsum}

%\documentclass[draft, twocolumn]{rps-esrl2020}

\def\papername{\jobname}

\def\ds{\displaystyle}

%\usepackage[authordate,backend=biber]{biblatex-chicago}


\begin{document}

\markboth{E. Miralles-Dolz, A. Gray, M. de Angelis, E. Patelli}{Interval-Based Sensitivity Analysis for Epistemic Uncertainty}

%%%%%%%%%%%%%%%%%%%%%%%%% Plase keep this command for single column for abstract section.
\twocolumn[
%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Interval-Based Sensitivity Analysis for Epistemic Uncertainty}

\author{Enrique Miralles-Dolz}

\address{Institute for Risk and Uncertainty, University of Liverpool, United Kingdom. \email{enmidol@liverpool.ac.uk}}
\address{Culham Centre for Fusion Energy, United Kingdom Atomic Energy Authority, United Kingdom}

\author{Ander Gray}

\address{Institute for Risk and Uncertainty, University of Liverpool, United Kingdom. \email{akgray@liverpool.ac.uk}}
\address{Culham Centre for Fusion Energy, United Kingdom Atomic Energy Authority, United Kingdom}

\author{Marco de Angelis}

\address{Institute for Risk and Uncertainty, University of Liverpool, United Kingdom. \email{mda@liverpool.ac.uk}}

\author{Edoardo Patelli}

\address{Centre for Intelligent Infrastructure, University of Strathclyde, United Kingdom. \email{edoardo.patelli@strath.ac.uk}}

\begin{abstract}
	The objective of sensitivity analysis is to understand how the input uncertainty of a mathematical model contributes to its output uncertainty.
	In the context of a digital twin, sensitivity analysis is paramount for the automatic verification and validation of physical models, and can also
	be used as a decision support tool to determine on which parameter to invest more empirical effort. Yet, sensitivity analysis often requires making
	assumptions about the inputs such as probability distribution functions, as it is the case in variance-based methods, or relies on surrogate models
	that also introduce more assumptions, as in kriging or polynomial chaos. It can be the case that one cannot reliably assign probability distribution
	functions if the model is dominated by epistemic uncertainties, or the complexity of the model is such that surrogate models cannot accurately capture
	its behaviour.

	We present a non-probabilistic sensitivity analysis method which requires no assumptions about the input probability distributions: the uncertainty in
	the input is expressed in the form of intervals, and employs the width of the output interval as the only measure, in the same way that interval analysis.
	As a positive by-product, the method also returns all the information that could be obtained with the reduction of the input interval to a single value,
	also called pinching. We use the Ishigami function as test case to show the performance of the proposed method, and compare it with Sobol indices.
\end{abstract}

\keywords{epistemic uncertainty, uncertainty quantification, sensitivity analysis, interval arithmetic, sobol indices, ishigami function.}

%%%%%%%%%%%%%%%%%%%%%%%%% Please keep this closing bracket to complete the single column format for abstract.
]
%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

Prediction is inherent in science, because prediction is essential to test theories and their consequences.
Thanks to the power of modern computation, the prediction of mathematical models (their output) representing natural phenomena can now be tested at unprecedented scales.
Digital twins attempt to exploit this advantage by modelling physical systems and allowing their evaluation via simulation, as these systems are often technically and
economically prohibitive to operate in the physical world.
However, digital twins are often so complex that it is not possible to infer their prediction from experience and judgement alone.
Therefore, it is necessary to verify and validate the prediction of these models without having to rely on subjectivity.
Sensitivity analysis can help with this task, by indicating what model parameters are responsible for the prediction of the model, and how
that prediction depends on them.

Generally, sensitivity analysis methods fall within three categories: derivative-based, distribution-based, and regression-based [REFerence future].
Derivative-based approaches attempt to compute the derivative of the model functions, either analytically or numerically, and measure the change in
the output when the inputs are perturbed around a base point [reference].
Distribution-based methods, such as Sobol' indices, decompose the output variance and assigns the partitions to the input variances, indicating how much
of the output variance is caused by each input variance [reference].
Lastly, regression-based approaches employ correlation coefficients, regression coefficients, or other machine learning methods [reference].

However, these approaches present some limitations that can make them unsuitable in certain cases.
For instance, the analytical description of the functions in the model are not always available, since it is not uncommon to deal with black-box models or
models with too many functions that make unpractical or difficult their analytical derivation.
Also, derivative-based methods require defining a base point for each input parameter, and a perturbation size.
It is not rare to find a situation where there is not consensus about those parameters.
A similar argument can be made for distribution-based methods, which require a precise definition of the probability distribution functions of the model input
parameters.
Lastly, to successfully apply regression-based methods, it is necessary to know the behavior of the model under investigation, which is not always known.
For example, correlation coefficients are suitable when the model is monotonic (or rank coefficients when it is non-monotonic).
For these reasons, it is desirable to find a sensitivity analysis method that:

\begin{enumerate}
	\item Does not require knowing the analytical description of the model functions.
	\item Does not require defining base values for any input parameter.
	\item Does not require to assume the input parameters follow a precise probability distribution function.
	\item It is independent of the model behavior.
\end{enumerate}

This paper attempts to present an interval-based sensitivity analysis method that fulfills these requirements.
%mention that 2,3,4 are related to epistemic uncertainty?
Section 2 introduces some basic concepts in interval arithmetic which are useful to introduce the method.
In Section 3 it is explained how the sensitivity indices are calculated in the interval approach, and the
pinching measure that can be calculated additionally.
Lastly, Section 4 compares the performance of the interval-based method against Sobol' indices in two test cases
(Ishigami function and a basic physics problem), followed by the conclusion in section 5.

\begin{enumerate}
	\item Explain sensitivity analysis. [DONE]
	\item Explain main approaches and their limitation. [DONE]
	\item Explain epistemic uncertainty in engineering -slides- (we cannot afford those limitations) [DONE]
	\item Explain structure of the paper [DONE]
\end{enumerate}

\section{Interval Arithmetic}

Minimum unit of assumption (or knowledge).

\begin{enumerate}
	\item Explain epistemic uncertainty can be modeled as intervals.
	\item Explain interval arithmetic propagates interval uncertainty.
	\item Explain issues with dependence, but there is some work done that could be useful.
	\item Explain that interval arithmetic requires knowledge of the analytical function, which is not
	always available. However, the same can be done with sampling, which does not require knowing the analytical function.
\end{enumerate}

\section{Interval-Based Sensitivity}

\begin{enumerate}
	\item Explain the area stuff
	\item Explain extreme cases (e.g. perfect dependence = index of 1 = no area, independence = index of 0 = all area)
	\item Explain pinching.
\end{enumerate}

\section{Application}

\begin{enumerate}
	\item Explain simple function (or physics based example?).
	\item Explain comparison with Sobol indices.
	\item Show results.
\end{enumerate}

\begin{enumerate}
	\item Explain Ishigami function
	\item Explain comparison with sobol indices (arguably the most common SA method)
	\item Show results.
\end{enumerate}


\section{Conclusion}


\begin{acknowledgement}
% Let's keep the aknowledgement in the same font as the body text
This section and its heading (unnumbered) should be in 9pt and come before the appendices and references.  Dedication and funding information may also be included here. 
\end{acknowledgement}

\section*{CItations and References}

Type references in continuous format in 9pt Times New Roman with single line spacing.   The references need to be in a proper style, i.e., the
Chicago bibliography style. Examples: \cite{lamport94,knuth79}. For more information, please visit: https://ctan.org/pkg/
% We should include an example bibfile with a couple of entries as well as citation in the body text
\bibliography{sample}
\bibliographystyle{Chicago}

\end{document}



